---
title: "Aboutreviews"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Aboutreviews}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## This page primarily serves as a proof-of-concept for certain website customization aspects using Pkgdown. One should not necessarily expect my website (at least this one) to be updated with my actual reviews.

Regrettably, all the review works presented here and all the rights reserved therein are the property of Ryan Ouillete (EIC) and "The VR Grid." This simply serves as a portfolio of the reviews I have completed in the past and the VR games I have experience with.

In my reviews, I aim to be as thorough as humanly possible, systematically documenting every possible pro and con that a given game or piece of equipment might have to offer. This might be off-putting, as they are typically by no means brief and delve into many subtle details. Still, I believe in the value of people's money and not supporting decisions to waste it. 

The VR games market is often flooded by short, niche "experiences" that contain the bare minimum of content outside of their mere medium. Since the release of blockbuster hits such as Boneworks and Half Life Alyx, the industry has clearly advanced to a point where this attitude will no longer suffice. Many of my reviews, the astute reader may notice, are exceptionally strict. This has warranted some controversy from game developers that see my reviews, but I will not yield.

After starting my graduate school career in cognitive neuroscience, and then cognitive and comparative psychology, I had to take an extended hiatus on my gaming reviewer journey. I am hoping, one day, I will be able to make a comeback and see what the industry has to offer in 2022. 

If anyone has any questions, comments, or concerns- or if any developers would like me to review their games- feel free to contact me at my email, danb.siegel@gmail.com.

Thanks for visiting!
```{r setup}
library(DBSStats2Labs)
```
