---
title: "Final_Lab_WYOR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Final_Lab_WYOR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Note: This is effectively a condensed version of all the functions used in my "immuno.analyze" package. 

### This function requires loading of the "stringr," "Tidyr," and "dplyr" packages.

### This function assumes that, when imaging your brain tissue, a group-specific tag is put at the beginning of the file name for each of the groups. The first of these will ultimately become the same thing input under the "first_group" term.

### It is recommended that groups are code-named for anonymity of condition to avoid experimenter bias, along all steps of the immunohistochemistry process, if possible.  

### See the "immuno.analyze" package for working versions of these functions. Documentation for this package is available at https://drseacow.github.io/immuno.analyze/ .

### Percent area analysis, averaging over coronal plane of tissue

overall.percent.area <- function(graph_name, DV, first_group, second_group, n_per_group, data = x){

  library(stringr)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(patchwork)

  immunot_data <- as.data.frame(read.csv(data, header = TRUE, sep = ",", dec = ".")) %>%
    select(c("Label", "Mean", "X.Area"))
  immunot_data <- separate(immunot_data, "Label", c("Group", NA, "Subject", NA, "Plane", "Side", NA, NA, NA, NA, NA, NA), "_")

  gp1_grand_mean <- mean(immunot_data[immunot_data$Group == first_group, 'X.Area'])
  gp2_grand_mean <- mean(immunot_data[immunot_data$Group == second_group, 'X.Area'])

  gp1_grand_sterr <- sd(immunot_data[immunot_data$Group == first_group, 'X.Area']) / sqrt(n_per_group)
  gp2_grand_sterr <- sd(immunot_data[immunot_data$Group == second_group, 'X.Area']) / sqrt(n_per_group)

  cross_plane_sterrs <- c(gp1_grand_sterr, gp2_grand_sterr)

  cross_plane_compare <- as.numeric(c(gp1_grand_mean, gp2_grand_mean))

  graph_df_without_plane <- tibble(Group = c(first_group, second_group),
                                   DV = cross_plane_compare)

  plot <- ggplot(graph_df_without_plane, aes(x = Group, y = DV, ylab = DV)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (DV - cross_plane_sterrs), ymax = (DV + cross_plane_sterrs))) +
    ggtitle(graph_name) +
    ylab(DV)

  gp1_grand_values <- c(immunot_data[immunot_data$Group == first_group, 'X.Area'])
  gp2_grand_values <- c(immunot_data[immunot_data$Group == second_group, 'X.Area'])

  overall_t <- t.test(gp1_grand_values, gp2_grand_values, var.equal = TRUE)

  analysis_display <- print(overall_t)
  analysis <- return(list(analysis_display, plot))

}

### Percent area analyses, differentiating between three coronal planes of tissue

percent.area.plane <- function(graph_name_1, graph_name_2, graph_name_3, DV, first_group, second_group, n_per_group, data = x){

  library(stringr)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(patchwork)

  immunot_data <- as.data.frame(read.csv(data, header = TRUE, sep = ",", dec = ".")) %>%
    select(c("Label", "Mean", "X.Area"))
  immunot_data <- separate(immunot_data, "Label", c("Group", NA, "Subject", NA, "Plane", "Side", NA, NA, NA, NA, NA, NA), "_")

  means_by_plane <- tibble(aggregate(immunot_data$X.Area, FUN = mean, by = list(Plane = immunot_data$Plane, Group = immunot_data$Group)))

  sds_by_plane <- tibble(aggregate(immunot_data$X.Area, FUN = sd, by = list(Plane = immunot_data$Plane, Group = immunot_data$Group)))

  gp1_ant <- immunot_data[immunot_data$Group == first_group & immunot_data$Plane == "Ant", ]
  aa <- as.vector(t(gp1_ant %>% select(X.Area)))
  gp2_ant <- immunot_data[immunot_data$Group == second_group & immunot_data$Plane == "Ant", ]
  ab <- as.vector(t(gp2_ant %>% select(X.Area)))

  gp1_med <- immunot_data[immunot_data$Group == first_group & immunot_data$Plane == "Med", ]
  ba <- as.vector(t(gp1_med %>% select(X.Area)))
  gp2_med <- immunot_data[immunot_data$Group == second_group & immunot_data$Plane == "Med", ]
  bb <- as.vector(t(gp2_med %>% select(X.Area)))

  gp1_pos <- immunot_data[immunot_data$Group == first_group & immunot_data$Plane == "Pos", ]
  ca <- as.vector(t(gp1_pos %>% select(X.Area)))
  gp2_pos <- immunot_data[immunot_data$Group == second_group & immunot_data$Plane == "Pos", ]
  cb <- as.vector(t(gp2_pos %>% select(X.Area)))

  gp1_mean_ant <- means_by_plane[1,3]
  gp1_mean_med <- means_by_plane[2,3]
  gp1_mean_pos <- means_by_plane[3,3]
  gp2_mean_ant <- means_by_plane[4,3]
  gp2_mean_med <- means_by_plane[5,3]
  gp2_mean_pos <- means_by_plane[6,3]

  gp1_sterr_ant <- sds_by_plane[1,3] / sqrt(n_per_group)
  gp1_sterr_med <- sds_by_plane[2,3] / sqrt(n_per_group)
  gp1_sterr_pos <- sds_by_plane[3,3] / sqrt(n_per_group)
  gp2_sterr_ant <- sds_by_plane[4,3] / sqrt(n_per_group)
  gp2_sterr_med <- sds_by_plane[5,3] / sqrt(n_per_group)
  gp2_sterr_pos <- sds_by_plane[6,3] / sqrt(n_per_group)

  ant_compare <- as.numeric(c(gp1_mean_ant, gp2_mean_ant))

  med_compare <- as.numeric(c(gp1_mean_med, gp2_mean_med))

  pos_compare <- as.numeric(c(gp1_mean_pos, gp2_mean_pos))

  ant_sterrs <- as.numeric(c(gp1_sterr_ant, gp2_sterr_ant))

  med_sterrs <- as.numeric(c(gp1_sterr_med, gp2_sterr_med))

  pos_sterrs <- as.numeric(c(gp1_sterr_pos, gp2_sterr_pos))

  graph_df_ant <- tibble(Ant = rep(c(first_group, second_group)),
                         DV = ant_compare)

  graph_df_med <- tibble(Med = rep(c(first_group, second_group)),
                         DV = med_compare)

  graph_df_pos <- tibble(Pos = rep(c(first_group, second_group)),
                         DV = pos_compare)

  a <- ggplot(graph_df_ant, aes(x = Ant, y = DV)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (DV - ant_sterrs), ymax = (DV + ant_sterrs))) +
    ggtitle(graph_name_1) +
    ylab(DV)
  m <- ggplot(graph_df_med, aes(x = Med, y = DV)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (DV - med_sterrs), ymax = (DV + med_sterrs))) +
    ggtitle(graph_name_2) +
    ylab(DV)
  p <- ggplot(graph_df_pos, aes(x = Pos, y = DV)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (DV - pos_sterrs), ymax = (DV + pos_sterrs))) +
    ggtitle(graph_name_3) +
    ylab(DV)

  ant_t <- t.test(aa, ab, var.equal = TRUE)$p.value
  med_t <- t.test(ba, bb, var.equal = TRUE)$p.value
  pos_t <- t.test(ca, cb, var.equal = TRUE)$p.value

  analysis_list <- c(ant_t, med_t, pos_t)
  plot <- a+m+p

  analysis <- return(list(analysis_list, plot))

}


### Cell count analyses, averaging over coronal plane of tissue

overall.cell.counts <- function(graph_name, DV, first_group, second_group, n_per_group, data = x){

  library(stringr)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(patchwork)

  immunot_data <- as.data.frame(read.csv(data, header = TRUE, sep = ",", dec = ".")) %>%
    select(c("Label", "Mean", "X.Area"))
  immunot_data <- separate(immunot_data, "Label", c("Group", NA, "Subject", NA, "Plane", "Side", NA), "_")

  values_by_group <- tibble(aggregate(immunot_data$Plane, FUN = length, by = list(Plane = immunot_data$Plane, Group = immunot_data$Group, immunot_data$Subject)))

  values_by_group_avg_by_plane <- tibble(aggregate(values_by_group$x, FUN = mean, by = list(Group = values_by_group$Group, Subject = values_by_group$Group.3)))

  gp1_values <- as.vector(values_by_group_avg_by_plane$x[values_by_group_avg_by_plane$Group == first_group])
  gp2_values <- as.vector(values_by_group_avg_by_plane$x[values_by_group_avg_by_plane$Group == second_group])

  gp1_grand_mean <- sum(values_by_group_avg_by_plane[values_by_group_avg_by_plane$Group == first_group, 'x']) / n_per_group
  gp2_grand_mean <- sum(values_by_group_avg_by_plane[values_by_group_avg_by_plane$Group == second_group, 'x']) / n_per_group

  gp1_sterr <- sd(gp1_values) / sqrt(n_per_group)
  gp2_sterr <- sd(gp2_values) / sqrt(n_per_group)

  cross_plane_sds <- c(gp1_sterr, gp2_sterr)

  cross_plane_compare <- as.numeric(c(gp1_grand_mean, gp2_grand_mean))

  graph_df_without_plane <- tibble(Group = c(first_group, second_group),
                                   DV = cross_plane_compare)

  plot <- ggplot(graph_df_without_plane, aes(x = Group, y = DV)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (DV - cross_plane_sds), ymax = (DV + cross_plane_sds))) +
    ggtitle(graph_name) +
    ylab(DV)

  overall_t <- t.test(gp1_values, gp2_values, var.equal = TRUE)
  overall_t

  analysis_display <- print(overall_t)
  return(plot)

}

### Cell count analyses, differentiating between three coronal planes of tissue

cell.counts.plane <- function(graph_name_1, graph_name_2, graph_name_3, DV, first_group, second_group, n_per_group, data = x){

  library(stringr)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(patchwork)

  immunot_data <- as.data.frame(read.csv(data, header = TRUE, sep = ",", dec = ".")) %>%
    select(c("Label", "Mean", "X.Area"))
  immunot_data <- separate(immunot_data, "Label", c("Group", NA, "Subject", NA, "Plane", "Side", NA), "_")

  counts_by_plane <- tibble(aggregate(immunot_data$Plane, FUN = length, by = list(Plane = immunot_data$Plane, Group = immunot_data$Group, immunot_data$Subject)))

  gp1_ant <- counts_by_plane[counts_by_plane$Group == first_group & counts_by_plane$Plane == "Ant", ]
  aa <- as.vector(t(gp1_ant %>% select(x)))
  gp2_ant <- counts_by_plane[counts_by_plane$Group == second_group & counts_by_plane$Plane == "Ant", ]
  ab <- as.vector(t(gp2_ant %>% select(x)))

  gp1_med <- counts_by_plane[counts_by_plane$Group == first_group & counts_by_plane$Plane == "Med", ]
  ba <- as.vector(t(gp1_med %>% select(x)))
  gp2_med <- counts_by_plane[counts_by_plane$Group == second_group & counts_by_plane$Plane == "Med", ]
  bb <- as.vector(t(gp2_med %>% select(x)))

  gp1_pos <- counts_by_plane[counts_by_plane$Group == first_group & counts_by_plane$Plane == "Pos", ]
  ca <- as.vector(t(gp1_pos %>% select(x)))
  gp2_pos <- counts_by_plane[counts_by_plane$Group == second_group & counts_by_plane$Plane == "Pos", ]
  cb <- as.vector(t(gp2_pos %>% select(x)))

  plane_count <- tibble(Group = rep(rep(c("gp1", "gp2"), each = 6), 3),
                        Plane = rep(c("ant", "med", "pos"), each = 12),
                        Counts = c(aa, ab, ba, bb, ca, cb)
  )

  means_by_plane <- aggregate(counts_by_plane$x, FUN = mean, by = list(Plane = counts_by_plane$Plane, Group = counts_by_plane$Group))

  sds_by_plane <- tibble(aggregate(counts_by_plane$x, FUN = sd, by = list(Plane = counts_by_plane$Plane, Group = counts_by_plane$Group)))

  gp1_mean_ant <- means_by_plane[1,3]
  gp1_mean_med <- means_by_plane[2,3]
  gp1_mean_pos <- means_by_plane[3,3]
  gp2_mean_ant <- means_by_plane[4,3]
  gp2_mean_med <- means_by_plane[5,3]
  gp2_mean_pos <- means_by_plane[6,3]

  gp1_sterr_ant <- as.numeric(sds_by_plane[1,3]) / sqrt(length(n_per_group))
  gp1_sterr_med <- as.numeric(sds_by_plane[2,3]) / sqrt(length(n_per_group))
  gp1_sterr_pos <- as.numeric(sds_by_plane[3,3]) / sqrt(length(n_per_group))
  gp2_sterr_ant <- as.numeric(sds_by_plane[4,3]) / sqrt(length(n_per_group))
  gp2_sterr_med <- as.numeric(sds_by_plane[5,3]) / sqrt(length(n_per_group))
  gp2_sterr_pos <- as.numeric(sds_by_plane[6,3]) / sqrt(length(n_per_group))

  ant_compare <- c(gp1_mean_ant, gp2_mean_ant)

  med_compare <- c(gp1_mean_med, gp2_mean_med)

  pos_compare <- c(gp1_mean_pos, gp2_mean_pos)

  ant_sterrs <- as.numeric(c(gp1_sterr_ant, gp2_sterr_ant))

  med_sterrs <- as.numeric(c(gp1_sterr_med, gp2_sterr_med))

  pos_sterrs <- as.numeric(c(gp1_sterr_pos, gp2_sterr_pos))

  graph_df_ant <- tibble(Ant = rep(c(first_group, second_group)),
                         DV = ant_compare)

  graph_df_med <- tibble(Med = rep(c(first_group, second_group)),
                         DV = med_compare)

  graph_df_pos <- tibble(Pos = rep(c(first_group, second_group)),
                         DV = pos_compare)

  a <- ggplot(graph_df_ant, aes(x = Ant, y = ant_compare)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (ant_compare - ant_sterrs), ymax = (ant_compare + ant_sterrs))) +
    ggtitle(graph_name_1) +
    ylab(DV)

  m <- ggplot(graph_df_med, aes(x = Med, y = DV)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (med_compare - med_sterrs), ymax = (med_compare + med_sterrs))) +
    ggtitle(graph_name_2) +
    ylab(DV)

  p <- ggplot(graph_df_pos, aes(x = Pos, y = DV)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (pos_compare - pos_sterrs), ymax = (pos_compare + pos_sterrs))) +
    ggtitle(graph_name_3) +
    ylab(DV)

  ant_t <- t.test(aa, ab, var.equal = TRUE)
  med_t <- t.test(ba, bb, var.equal = TRUE)
  pos_t <- t.test(ca, cb, var.equal = TRUE)

  analysis_list <- c(ant_t$p.value, med_t$p.value, pos_t$p.value)

  plot <- a+m+p

  return(list(analysis_list, plot))

}


### Mean cell intensity analyses, averging over coronal plane of tissue

overall.intensity <- function(graph_name, DV, first_group, second_group, n_per_group, data = x){

  immunot_data <- as.data.frame(read.csv(data, header = TRUE, sep = ",", dec = ".")) %>%
    select(c("Label", "Mean", "X.Area"))
  immunot_data <- separate(immunot_data, "Label", c("Group", NA, "Subject", NA, "Plane", "Side", NA), "_")

  gp1_grand_mean <- mean(immunot_data[immunot_data$Group == first_group, 'Mean'])
  gp2_grand_mean <- mean(immunot_data[immunot_data$Group == second_group, 'Mean'])

  gp1_grand_sterr <- sd(immunot_data[immunot_data$Group == first_group, 'Mean']) / sqrt(n_per_group)
  gp2_grand_sterr <- sd(immunot_data[immunot_data$Group == second_group, 'Mean']) / sqrt(n_per_group)

  cross_plane_sterr <- c(gp1_grand_sterr, gp2_grand_sterr)

  cross_plane_compare <- as.numeric(c(gp1_grand_mean, gp2_grand_mean))

  graph_df_without_plane <- tibble(Group = c(first_group, second_group),
                                   DV = cross_plane_compare)

  plot <- ggplot(graph_df_without_plane, aes(x = Group, y = cross_plane_compare)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (cross_plane_compare - cross_plane_sterr), ymax = (cross_plane_compare + cross_plane_sterr))) +
    ggtitle(graph_name) +
    ylab(DV)

  values_by_group <- tibble(aggregate(immunot_data$Mean, FUN = mean, by = list(Group = immunot_data$Group, Subject = immunot_data$Subject)))

  values_by_group_avg_by_plane <- tibble(aggregate(values_by_group$x, FUN = mean, by = list(Group = values_by_group$Group, Subject = values_by_group$Subject)))

  gp1_values <- as.vector(values_by_group$x[values_by_group$Group == first_group])
  gp2_values <- as.vector(values_by_group$x[values_by_group$Group == second_group])

  overall_t <- t.test(gp1_values, gp2_values, var.equal = TRUE)
  return(list(overall_t, plot))

}


### Mean cell intensity analyses, differentiating between three coronal planes of tissue

intensity.plane <- function(graph_name_1, graph_name_2, graph_name_3, DV, first_group, second_group, n_per_group, data = x){

  immunot_data <- as.data.frame(read.csv(data, header = TRUE, sep = ",", dec = ".")) %>%
    select(c("Label", "Mean", "X.Area"))
  immunot_data <- separate(immunot_data, "Label", c("Group", NA, "Subject", NA, "Plane", "Side", NA), "_")

  means_by_plane <- tibble(aggregate(immunot_data$Mean, FUN = mean, by = list(Plane = immunot_data$Plane, Group = immunot_data$Group)))

  sds_by_plane <- tibble(aggregate(immunot_data$Mean, FUN = sd, by = list(Plane = immunot_data$Plane, Group = immunot_data$Group)))

  values_by_plane <- tibble(aggregate(immunot_data$Mean, FUN = mean, by = list(Plane = immunot_data$Plane, Group = immunot_data$Group, Subject = immunot_data$Subject)))

  gp1_ant <-   values_by_plane[values_by_plane$Group == first_group & values_by_plane$Plane == "Ant", ]
  aa <- as.vector(t(gp1_ant %>% select(x)))
  gp2_ant <-   values_by_plane[values_by_plane$Group == second_group & values_by_plane$Plane == "Ant", ]
  ab <- as.vector(t(gp2_ant %>% select(x)))

  gp1_med <-   values_by_plane[values_by_plane$Group == first_group & values_by_plane$Plane == "Med", ]
  ba <- as.vector(t(gp1_med %>% select(x)))
  gp2_med <-   values_by_plane[values_by_plane$Group == second_group &  values_by_plane$Plane == "Med", ]
  bb <- as.vector(t(gp2_med %>% select(x)))

  gp1_pos <-   values_by_plane[values_by_plane$Group == first_group & values_by_plane$Plane == "Pos", ]
  ca <- as.vector(t(gp1_pos %>% select(x)))
  gp2_pos <-   values_by_plane[values_by_plane$Group == second_group &  values_by_plane$Plane == "Pos", ]
  cb <- as.vector(t(gp2_pos %>% select(x)))

  gp1_mean_ant <- means_by_plane[1,3]
  gp1_mean_med <- means_by_plane[2,3]
  gp1_mean_pos <- means_by_plane[3,3]
  gp2_mean_ant <- means_by_plane[4,3]
  gp2_mean_med <- means_by_plane[5,3]
  gp2_mean_pos <- means_by_plane[6,3]

  gp1_sterr_ant <- sds_by_plane[1,3] / sqrt(n_per_group)
  gp1_sterr_med <- sds_by_plane[2,3] / sqrt(n_per_group)
  gp1_sterr_pos <- sds_by_plane[3,3] / sqrt(n_per_group)
  gp2_sterr_ant <- sds_by_plane[4,3] / sqrt(n_per_group)
  gp2_sterr_med <- sds_by_plane[5,3] / sqrt(n_per_group)
  gp2_sterr_pos <- sds_by_plane[6,3] / sqrt(n_per_group)

  ant_compare <- as.numeric(c(gp1_mean_ant, gp2_mean_ant))

  med_compare <- as.numeric(c(gp1_mean_med, gp2_mean_med))

  pos_compare <- as.numeric(c(gp1_mean_pos, gp2_mean_pos))

  ant_sterrs <- as.numeric(c(gp1_sterr_ant, gp2_sterr_ant))

  med_sterrs <- as.numeric(c(gp1_sterr_med, gp2_sterr_med))

  pos_sterrs <- as.numeric(c(gp1_sterr_pos, gp2_sterr_pos))

  graph_df_ant <- tibble(Ant = rep(c(first_group, second_group)),
                         DV = ant_compare)

  graph_df_med <- tibble(Med = rep(c(first_group, second_group)),
                         DV = med_compare)

  graph_df_pos <- tibble(Pos = rep(c(first_group, second_group)),
                         DV = pos_compare)

  a <- ggplot(graph_df_ant, aes(x = Ant, y = ant_compare)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (ant_compare - ant_sterrs), ymax = (ant_compare + ant_sterrs))) +
    ggtitle(graph_name_1) +
    ylab(DV)

  m <- ggplot(graph_df_med, aes(x = Med, y = med_compare)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (med_compare - med_sterrs), ymax = (med_compare + med_sterrs))) +
    ggtitle(graph_name_2) +
    ylab(DV)

  p <- ggplot(graph_df_pos, aes(x = Pos, y = pos_compare)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = (pos_compare - pos_sterrs), ymax = (pos_compare + pos_sterrs))) +
    ggtitle(graph_name_3) +
    ylab(DV)

  ant_t <- t.test(aa, ab, var.equal = TRUE)
  med_t <- t.test(ba, bb, var.equal = TRUE)
  pos_t <- t.test(ca, cb, var.equal = TRUE)

  analysis_list <- c(ant_t$p.value, med_t$p.value, pos_t$p.value)

  plot <- a+m+p

  return(list(analysis_list, plot))

}

```{r setup}
library(DBSStats2Labs)
```
