---
title: "Lab4"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lab4}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Generalization Assignment

### Q1: Run and prove equivalence of ANOVA and T-Test on following data

```{r}

library(tibble)
library(dplyr)


example_data <- tibble(Group = rep(c("A", "B"), each = 5),
                       DV = c(2, 4, 3, 5, 4, 7, 6, 5, 6, 7))

## The t-test

Ttest <- t.test(DV ~ Group, var.equal = TRUE, data = example_data)
Ttest

```

```{r}
## ANOVA

example_data$DV <- as.vector(example_data$DV)

grand_mean <- mean(example_data$DV)

### SS Total

SS_totals <- example_data %>%
  mutate(grand_mean = mean(example_data$DV)) %>%
  mutate(deviations = DV - grand_mean,
         sq_deviations = (DV - grand_mean)^2)

SS_total <- sum(SS_totals$sq_deviations)

group_means <- example_data %>%
  group_by(Group) %>%
  summarize(mean_DV = mean(DV), .groups = 'drop')
  
### SS Between 

SS_betweens <- example_data %>%
  mutate(grand_mean = mean(example_data$DV),
         group_means = rep(group_means$mean_DV, each = 5)) %>%
  mutate(deviations = group_means - grand_mean,
         sq_deviations = (group_means - grand_mean)^2)

SS_between <- sum(SS_betweens$sq_deviations)

### SS Within

group_means <- example_data %>%
  group_by(Group) %>%
  summarize(mean_DV = mean(DV), .groups = 'drop')

SS_withins <- example_data %>%
  mutate(group_means = rep(group_means$mean_DV, each = 5)) %>%
  mutate(deviations = group_means - DV,
         sq_deviations = (group_means - DV)^2)

SS_within <- sum(SS_withins$sq_deviations)

SS_total

SS_between + SS_within

SS_total == SS_between + SS_within

#### Despite appearing to have the same values, these SSs are different. I'll proceed, but perhaps there's an issue with very small digits and rounding?

dfb <- 2 - 1
MS_Between <- SS_between / dfb

dfw <- 10 - 2
MS_Within <- SS_within / dfw

F_ratio <- MS_Between / MS_Within
F_ratio
### F is 16.9
print(pf(16.9, 1, 8, lower.tail = FALSE))

### P-value is 0.003386143; reject null. Does the automated method agree?

Aov <- summary(aov(DV ~ Group, data = example_data))
Aov

```

```{r}

###Checking whether the results (P-Value) agree:

t.test(DV ~ Group, var.equal = TRUE, data = example_data)$p.value == pf(16.9, 1, 8, lower.tail = FALSE)
```

```{r}
round(Aov[[1]]$`F value`[1]) == round((Ttest$statistic)^2)
```

### For this question 1, I was able to computationally solve for the ANOVA (not using the function) with slight reference to the lab project. Due to slight confusions, and being a bit too desperate to manually run the T-Test despite being confused on how to select each group mean, I needed veeeery slight reminders from the video on running these test functions. I was able to prove the p-values are the same without help, and in theory would've known how to prove the relation between P and T, but this latter bit needed some video consultation regarding the grammar of the situation.

### Q2 Reference to your undergraduate class ANOVA
```{r}

library(data.table)
anova_lab_data <- fread("https://raw.githubusercontent.com/CrumpLab/statisticsLab/master/data/Jamesetal2015Experiment2.csv")

library(ggplot2)

anova_lab_data

```

```{r}

anova_lab_data$Condition <- as.factor(anova_lab_data$Condition)
levels(anova_lab_data$Condition) <- c("Control",
                                      "Reactivation+Tetris",
                                      "Tetris_Only",
                                      "Reactivation_Only")

anova_lab_df <- anova_lab_data %>% 
  group_by(Condition) %>% 
  summarise(means = mean(Days_One_to_Seven_Number_of_Intrusions),
            SEs = sd(Days_One_to_Seven_Number_of_Intrusions) / sqrt(length(Days_One_to_Seven_Number_of_Intrusions)))

ggplot(anova_lab_df, aes(x = Condition, y = means)) + 
  geom_bar(stat = "identity", aes(fill = Condition)) + 
  geom_errorbar(aes(ymin = means - SEs,               
                    ymax = means + SEs), width = .1) +
  geom_point(data = anova_lab_data, aes(x = Condition, y=Days_One_to_Seven_Number_of_Intrusions), alpha = .5) +
  geom_point(alpha = .25) +
  ylab("Intrusive Memories (Mean for Week)")

```

```{r}

anova_lab_data_anova <- summary(aov(Days_One_to_Seven_Number_of_Intrusions ~ Condition, anova_lab_data))

anova_lab_data_anova

```
### For some reason, also, Papaja does not show up in my package installation options and I cannot figure it out. Nevertheless:

## One-way analysis of variance revealed an effect of experimental condition on the number of intrusive memories reported over the span of a week, F[3, 68] = 3.795, MSE = 10.09, P = 0.0141.

### Help-report: For this second question, I wasn't exactly sure what was required (as far as what variables, what aspects of the graph, etc) so I thoroughly consulted the original lab project documentation in which this data is linked. I had numerous super-bizarre erroneous moments with the wrong number of levels mysteriously being loaded, so I ended up consulting the solutions video, but I don't think I really need to. Again, humbly, I was unable to install Papaja but I hope my APA results sentence is at least somewhat sufficient.

## Follow-along with exercises

```{r}

library(tibble)
romeo_juliet <- tibble(subjects = 1:20,
                       Group = rep(c("No Context",
                                 "Context Before",
                                 "Context After",
                                 "Partial Context"), each = 5),
                       Comprehension = c(3,3,2,4,3,
                                         5,9,8,4,9,
                                         2,4,5,4,1,
                                         5,4,3,5,4
                                   )
                          )

romeo_juliet$Group <- factor(romeo_juliet$Group,
                             levels = c("No Context",
                                 "Context Before",
                                 "Context After",
                                 "Partial Context"))

knitr::kable(romeo_juliet)

```

### GrandMean and Total SS

```{r}
library(dplyr)
grand_mean <- mean(romeo_juliet$Comprehension)

SS_total_table <- romeo_juliet %>% 
  mutate(grand_mean = mean(romeo_juliet$Comprehension)) %>% 
  mutate(deviations = Comprehension - grand_mean,
         sq_deviations = (Comprehension - grand_mean)^2)

SS_total <- sum(SS_total_table$sq_deviations)
```

### Between SS

```{r}

group_means <- romeo_juliet %>% 
  group_by(Group)%>%
  summarize(mean_Comprehension = mean(Comprehension), .groups = 'drop')

SS_between_table <- romeo_juliet %>% 
  mutate(grand_mean = mean(romeo_juliet$Comprehension),
         group_means = rep(group_means$mean_Comprehension, each = 5)) %>%
  mutate(deviations = group_means - grand_mean,
         sq_deviations = (group_means - grand_mean)^2)

SS_between <- sum(SS_between_table$sq_deviations)
  

```

### Within SS

```{r}

group_means <- romeo_juliet %>% 
  group_by(Group) %>%
  summarize(mean_Comprehension = mean(Comprehension), .groups = 'drop')

SS_within_table <- romeo_juliet %>% 
  mutate(group_means = rep(group_means$mean_Comprehension, each = 5)) %>%
  mutate(deviations = group_means - Comprehension,
         sq_deviations = (group_means - Comprehension)^2)

SS_within <- sum(SS_within_table$sq_deviations)
  

```

###Check of additivity

```{r}
SS_total

SS_between + SS_within

SS_total == SS_between + SS_within 
```

### F as a ratio of variances

```{r}

dfb <- 4 - 1
MS_Between <- SS_between / dfb

dfw <- 20 - 4
MS_Within <- SS_within / dfw

F_ratio <- MS_Between / MS_Within

```

### ANOVA Method using Matrices

```{r}

matrix_data <- matrix(c(3, 3, 2, 4, 3, 
                        5, 9, 8, 4, 9,
                        2, 4, 5, 4, 1,
                        5, 4, 3, 5, 4),
                      ncol = 4,
                      nrow = 5)

colnames(matrix_data) <- c("No Context",
                                "Context Before",
                                "Context After",
                                "Partial Context")

SS_total <- sum( (matrix_data - mean(matrix_data))^2 )
SS_between <- sum( (colMeans(matrix_data) - mean(matrix_data))^2 )*5
SS_within <- sum( (colMeans(matrix_data) - t(matrix_data))^2 )

dfb <- 4 - 1 
MS_Between <- SS_between / dfb

dfw <- 20 - 4
MS_Within <- SS_within / dfw

F_ratio <- MS_Between / MS_Within
```

### Using the "AOV" function

```{r}

romeo_juliet$Comprehension <- sample(romeo_juliet$Comprehension)

anova.out <- aov(Comprehension ~ Group, data = romeo_juliet)

summary(anova.out)

```

## Simulating Components of the ANOVA Table

### SS Total

```{r}

sim_data <- matrix(rnorm(20, 0, 1), ncol = 4, nrow = 5)

SS_total <- sum( (mean(sim_data) - sim_data)^2 )
SS_total

SS_total_distribution <- c()
for(i in 1:1000){
  sim_data <- matrix(rnorm(20, 0, 1), ncol = 4, nrow = 5)
  SS_total <- sum( (mean(sim_data) - sim_data)^2 )
  SS_total_distribution[i] <- SS_total
}
hist(SS_total_distribution)

mean(SS_total_distribution)

SS_total_distribution_alt <- c()
for(i in 1:1000){
  sim_data <- matrix(rnorm(20, 0, 1), ncol = 4, nrow = 5)
  sim_data[,1] <- sim_data[,1] + 2
  SS_total <- sum( (mean(sim_data) - sim_data)^2 )
  SS_total_distribution_alt[i] <- SS_total
}
hist(SS_total_distribution_alt)
mean(SS_total_distribution_alt)

library(ggplot2)
SS_total_data <- data.frame(SS_total = c(SS_total_distribution,
                                         SS_total_distribution_alt),
                            type = rep(c("Null", "Alternative"), each = 1000))
ggplot(SS_total_data, aes(x = SS_total, group = type, fill = type)) +
  geom_histogram(position = "dodge")

```

### Simulating SS Between

```{r}

SS_between_distribution <- c()
for(i in 1:1000){
  sim_data <- matrix(rnorm(20, 0, 1), ncol = 4, nrow = 5)
  SS_between <- sum( (mean(sim_data) - colMeans(sim_data))^2 )*5
  SS_between_distribution[i] <- SS_between
}

SS_between_distribution_alt <- c()
for(i in 1:1000){
  sim_data <- matrix(rnorm(20, 0, 1), ncol = 4, nrow = 5)
  sim_data[,1] <- sim_data[,1] + 2
  SS_between <- sum( (mean(sim_data) - colMeans(sim_data))^2 ) * 5
  SS_between_distribution_alt[i] <- SS_between
}

SS_between_data <- data.frame(SS_between = c(SS_between_distribution,
                                             SS_between_distribution_alt),
                              type = rep(c("Null", "Alternative"), each = 1000))
ggplot(SS_between_data, aes(x = SS_between, group = type, fill = type)) + 
  geom_histogram(position = "dodge")

```

### Simulating the SS Within

```{r}

SS_Within_distribution <- c()
for(i in 1:1000){
  sim_data <- matrix(rnorm(20, 0, 1), ncol = 4, nrow = 5)
  SS_Within <- sum( (colMeans(sim_data) - t(sim_data))^2 )
  SS_Within_distribution[i] <- SS_Within
}

SS_Within_distribution_alt <- c()
for(i in 1:1000){
  sim_data <- matrix(rnorm(20, 0, 1), ncol = 4, nrow = 5)
  sim_data[,1] <- sim_data[,1] + 2
  SS_Within <- sum( (colMeans(sim_data) - t(sim_data))^2)
  SS_Within_distribution_alt[i] <- SS_Within
}

SS_Within_data <- data.frame(SS_Within = c(SS_Within_distribution,
                                           SS_Within_distribution_alt),
                             type = rep(c("Null", "Alternative"), each = 1000))
ggplot(SS_Within_data, aes(x = SS_Within, group = type, fill = type)) +
  geom_histogram(position = "dodge")
```

### Simulating MS Between, MS Within, and the Resulting F-Ratio

```{r}

MS_between_data <- data.frame(MS_between = c(SS_between_distribution / 3,
                                             SS_between_distribution_alt / 3),
                              type = rep(c("Null", "Alternative"), each = 1000))
ggplot(MS_between_data, aes(x = MS_between, group = type, fill = type)) +
  geom_histogram(position = "dodge")

MS_Within_data <- data.frame(MS_Within = c(SS_Within_distribution / 16,
                                           SS_Within_distribution_alt / 16),
                             type = rep(c("Null", "Alternative"), each = 1000))
ggplot(MS_Within_data, aes(x = MS_Within, group = type, fill = type)) +
  geom_histogram(position = "dodge")

F_distribution <- c()
for(i in 1:1000){
  sim_data <- matrix(rnorm(20, 0, 1), ncol = 4, nrow = 5)
  SS_between <- sum( (mean(sim_data) - colMeans(sim_data))^2 ) * 5
  SS_Within <- sum( (colMeans(sim_data) - t(sim_data))^2 )
  sim_F <- (SS_between / 3) / (SS_Within / 16)
  F_distribution[i] <- sim_F
}

F_distribution_alt <- c()
for(i in 1:1000){
  sim_data <- matrix(rnorm(20, 0, 1), ncol = 4, nrow = 5)
  sim_data[,1] <- sim_data[,1]+2
  SS_between <- sum( (mean(sim_data) - colMeans(sim_data))^2 ) * 5
  SS_Within <- sum( (colMeans(sim_data) - t(sim_data))^2 )
  sim_F <- (SS_between / 3) / (SS_Within / 16)
  F_distribution_alt[i] <- sim_F
}

F_data <- data.frame(F = c(F_distribution,
                           F_distribution_alt),
                     type = rep(c("Null", "Alternative"),
                                each = 1000))
ggplot(F_data, aes(x = F, group = type, fill = type)) +
  geom_histogram(position = "dodge")
```

## Truly Simulating the F-Distribution Itself

```{r}

pf(7.227, 3, 16, lower.tail = FALSE)

romeo_juliet$Comprehension <- rnorm(20, 0, 1)
aov.out <- aov(Comprehension ~ Group, data = romeo_juliet)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]

save_F_values <- length(10000)
for(i in 1:10000){
  romeo_juliet$Comprehension <- rnorm(20, 0, 1)
  aov.out <- aov(Comprehension ~ Group, data = romeo_juliet)
  simulated_F <- summary(aov.out)[[1]]$`F value`[1]
  save_F_values[i] <- simulated_F
}

hist(save_F_values)

length(save_F_values[save_F_values > 7.22]) / length(save_F_values)
```

## One-Way ANOVAs Using the AOV Function

```{r}

romeo_juliet <- tibble(subjects = 1:20,
                       Group = rep(c("No Context",
                                     "Context Before",
                                     "Context After",
                                     "Partial Context"), each = 5),
                       Comprehension = c(3, 3, 2, 4, 3,
                                         5, 9, 8, 4, 9,
                                         2, 4, 5, 4, 1,
                                         5, 4, 3, 5, 4
                                    )
                            )

romeo_juliet$Group <- factor(romeo_juliet$Group,
                             levels = c("No Context",
                                        "Context Before",
                                        "Context After",
                                        "Partial Context"))

anova.out <- aov(Comprehension ~ Group, data = romeo_juliet)
anova.out
summary(anova.out)

#alternatively

summary(aov(Comprehension ~ Group, data = romeo_juliet))

model.tables(anova.out)

anova.out <- aov(Comprehension ~ Group, data = romeo_juliet)
summary(anova.out)
model.tables(anova.out)

my_summary <- summary(anova.out)
my_summary[[1]]$Df
my_summary[[1]]$`Sum Sq`
my_summary[[1]]$`Mean Sq`
my_summary[[1]]$`F value`
my_summary[[1]]$`Pr(>F)`

#Papaja doesn't seem to exist for me? I'm confused on how to get it running...

```



```{r setup}
library(DBSStats2SemesterProject)
```
