---
title: "Lab5a"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lab5a}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Generalization Assignment

### Q1 - Simulated Data with an F smaller than the critical value

```{r}
library(tibble)

levels <- 3
n_per_level <- 10

F_crit <- qf(0.95, 2, 27)
F_crit

for(i in 1:1000){
random_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = rnorm(levels*n_per_level, 0, 1)
                      )

F_crit <- qf(0.95, 2, 27)
F_crit

aov.out <- aov(DV ~ IV, data = random_data)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]
if(simulated_F < F_crit) break
}

summary(aov.out)
```

```{r}
library(ggplot2)

ggplot(random_data, aes(x = IV, y = DV))+
  geom_bar(stat = "summary", fun = "mean")+
  geom_point()
```

```{r}
i
```

### We would fail to reject the null hypothesis in this scenario, as- by design- the simulated F is below the critical F. Given the way this simulation is structured, we can be at least 95% confident (although this is a somewhat loaded way to phrase it) that this was the correct decision.

### Q2 - Simulated data that is above the critical value
```{r}
library(tibble)

levels <- 3
n_per_level <- 10

F_crit <- qf(0.95, 2, 27)
F_crit

for(i in 1:1000){
random_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = rnorm(levels*n_per_level, 0, 1)
                      )

F_crit <- qf(0.95, 2, 27)
F_crit

aov.out <- aov(DV ~ IV, data = random_data)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]
if(simulated_F > F_crit) break
}

summary(aov.out)
```
```{r}
i
```
### By design, we would reject the null hypothesis as the simulated F value here is greater than the critical F value. We would have 95% certainty that this was the correct decision. However, we know for a fact that this falls in the unfortunate 5% error range, as the data is randomly simulated and we have simply designed it to tell us whenever it gives us a lucky fraudulent F (so it would be the incorrect decision).

```{r}
library(ggplot2)

ggplot(random_data, aes(x = IV, y = DV))+
  geom_bar(stat = "summary", fun = "mean")+
  geom_point()
```

### Q3 - "Breaking" the assumptions of the normal distribution in the ANOVA

```{r}
save_F_values <- length(1000)
for(i in 1:1000){
  random_data <- tibble(subjects = 1:(levels*n_per_level),
                        IV = as.factor(rep(1:levels, each = n_per_level)),
                        DV = rnorm(levels*n_per_level) / rt(levels*n_per_level, 1, 1)
                        )
  aov.out <- aov(DV ~ IV, data = random_data)
  simulated_F <- summary(aov.out)[[1]]$`F value`[1]
  save_F_values[i] <- simulated_F
}

F_comparison <- tibble(type = rep(c("analytic","simulated"), each = 1000), 
                        F_value = c(rf(1000, levels - 1, (levels*n_per_level)),save_F_values))

ggplot(F_comparison, aes(x=F_value, color = type))+
  geom_freqpoly(bins = 50)
```

## Struggle-report: I'm sorry Matt, I admittedly needed profuse consultation of your solutions video on this one. I tried to do something a little different with Q3 and I basically did, although I followed the suggestion regarding the T-distribution. For Q1 and 2, I figured out what I needed to copy/paste and modify but couldn't figure out how to modify it (at least, without breaking the whole thing) without consulting the video. 

## Follow-along with class exercise

```{r}
someFs <- rf(1000,3,16)
hist(someFs)
```

```{r}
pf(4, 3, 16, lower.tail = FALSE)
```

```{r}

library(tibble)

levels <- 4
n_per_level <- 5

random_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = rnorm(levels*n_per_level, 0, 1)
                      )

aov.out <- aov(DV ~ IV, data = random_data)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]

save_F_values <- length(1000)
for(i in 1:1000){
random_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = rnorm(levels*n_per_level, 0, 1)
                      )
aov.out <- aov(DV ~ IV, data = random_data)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]
save_F_values[i] <- simulated_F
}

```

```{r}
library(ggplot2)

F_comparison <- tibble(type = rep(c("analytic","simulated"), each = 1000), 
                        F_value = c(rf(1000,3,16),save_F_values))

ggplot(F_comparison, aes(x=F_value))+
  geom_histogram()+
  facet_wrap(~type)
```

```{r}
save_new_F_values <- length(1000)
for(i in 1:1000){
random_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = rnorm(levels*n_per_level, 50, 25)
                      )
aov.out <- aov(DV ~ IV, data = random_data)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]
save_new_F_values[i] <- simulated_F
}

F_comparison <- tibble(type = rep(c("analytic","sim_0_1","sim_50_25"), each = 1000), 
                        F_value = c(rf(1000,3,16),save_F_values, save_new_F_values))

ggplot(F_comparison, aes(x=F_value))+
  geom_histogram()+
  facet_wrap(~type)
```

```{r}
hist(sample(c(rnorm(levels*n_per_level/2,0,1),rnorm(levels*n_per_level/2,5,1))))
```

```{r}

save_bimodal_F_values <- length(1000)
for(i in 1:1000){

  random_data <- tibble(subjects = 1:(levels*n_per_level),
                        IV = as.factor(rep(1:levels, each = n_per_level)),
                        DV = sample(c(rnorm(levels*n_per_level/2,0,1),
                                      rnorm(levels*n_per_level/2,5,1)))
                        )
  aov.out <- aov(DV ~ IV, data = random_data)
  simulated_F <- summary(aov.out)[[1]]$`F value`[1]
  save_bimodal_F_values[i] <- simulated_F
}

F_comparison <- tibble(type = rep(c("analytic","sim_0_1","sim_50_25", "sim_bimodal"), each = 1000), 
                        F_value = c(rf(1000,3,16),save_F_values, save_new_F_values,save_bimodal_F_values))

ggplot(F_comparison, aes(x=F_value))+
  geom_histogram(bins=100)+
  facet_wrap(~type)
```

```{r}
qf(.95,3,16)
```

```{r}
library(dplyr)

Fs <- F_comparison[F_comparison$type=="analytic",]$F_value
sorted_Fs <- sort(Fs)
sorted_Fs[950]

Fs <- F_comparison[F_comparison$type=="sim_0_1",]$F_value
sorted_Fs <- sort(Fs)
sorted_Fs[950]

Fs <- F_comparison[F_comparison$type=="sim_50_25",]$F_value
sorted_Fs <- sort(Fs)
sorted_Fs[950]

Fs <- F_comparison[F_comparison$type=="sim_bimodal",]$F_value
sorted_Fs <- sort(Fs)
sorted_Fs[950]
```

```{r}
levels <- 4
n_per_level <- 5

some_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = c(11,12,11,11,12,
                             10,8,10,9,10,
                             8,9,10,10,10,
                             10,8,10,9,10
                             )
                      )

aov.out <- aov(DV ~ IV, data = some_data)
summary(aov.out)
```

```{r}
save_F_values <- length(1000)
for(i in 1:1000){
  some_data <- tibble(subjects = 1:(levels*n_per_level),
                        IV = as.factor(rep(1:levels, each = n_per_level)),
                        DV = sample(c(11,12,11,11,12,
                               10,8,10,9,10,
                               8,9,10,10,10,
                               10,8,10,9,10
                               ))
                        )
  aov.out <- aov(DV ~ IV, data =some_data)
  simulated_F <- summary(aov.out)[[1]]$`F value`[1]
  save_F_values[i] <- simulated_F
}

F_comparison <- tibble(type = rep(c("analytic","randomization"), each = 1000), 
                        F_value = c(rf(1000,3,16),save_F_values))

ggplot(F_comparison, aes(x=F_value))+
  geom_histogram()+
  facet_wrap(~type)
```

```{r}
length(save_F_values[save_F_values>7.407])/1000
```
## Simulating the alternative hypothesis
```{r}

levels <- 4
n_per_level <- 5

alternative_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = c(rnorm(n_per_level, 125, 25),
                             rnorm(n_per_level, 100, 25),
                             rnorm(n_per_level, 100, 25),
                             rnorm(n_per_level, 100, 25)
                             )
                      )

save_altF_values <- length(1000)
for(i in 1:1000){
alternative_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = c(rnorm(n_per_level, 125, 25),
                             rnorm(n_per_level, 100, 25),
                             rnorm(n_per_level, 100, 25),
                             rnorm(n_per_level, 100, 25)
                             )
                      )
aov.out <- aov(DV ~ IV, data = alternative_data)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]
save_altF_values[i] <- simulated_F
}

F_comparison <- tibble(type = rep(c("null","alternative"), each = 1000), 
                        F_value = c(rf(1000,3,16),save_altF_values))

ggplot(F_comparison, aes(x=F_value))+
  geom_histogram(bins=100)+
  facet_wrap(~type, nrow=2)
```

```{r}
qf(.95,3,16)
```

```{r}
length(save_altF_values[save_altF_values > 3.23])
```

```{r}
n_per_level <- 50

# repeat the above many times to compute the F-distribution
save_altF_values <- length(1000)
for(i in 1:1000){
alternative_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = c(rnorm(n_per_level, 125, 25),
                             rnorm(n_per_level, 100, 25),
                             rnorm(n_per_level, 100, 25),
                             rnorm(n_per_level, 100, 25)
                             )
                      )
aov.out <- aov(DV ~ IV, data = alternative_data)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]
save_altF_values[i] <- simulated_F
}

F_comparison <- tibble(type = rep(c("null","alternative"), each = 1000), 
                        F_value = c(rf(1000,3,96),save_altF_values))

ggplot(F_comparison, aes(x=F_value))+
  geom_histogram(bins=100)+
  facet_wrap(~type, nrow=2)
```

```{r}
qf(.95,3,200-4)
```

```{r}
length(save_altF_values[save_altF_values > 2.65])
```

```{r}

n_per_level <- 25

# repeat the above many times to compute the F-distribution
save_altF_values <- length(1000)
for(i in 1:1000){
  alternative_data <- tibble(subjects = 1:(levels*n_per_level),
                        IV = as.factor(rep(1:levels, each = n_per_level)),
                        DV = c(rnorm(n_per_level, 125, 25),
                               rnorm(n_per_level, 100, 25),
                               rnorm(n_per_level, 100, 25),
                               rnorm(n_per_level, 100, 25)
                               )
                        )
  aov.out <- aov(DV ~ IV, data = alternative_data)
  simulated_F <- summary(aov.out)[[1]]$`F value`[1]
  save_altF_values[i] <- simulated_F
}

length(save_altF_values[save_altF_values > qf(.95,3,(n_per_level*4)-4)])/1000
```

```{r}

num_subjects <- c(5,10,15,20,25,30)
sim_power <- length(length(num_subjects))

for(n in 1:length(num_subjects)){

  n_per_level <- num_subjects[n]
  
  save_altF_values <- length(1000)
  for(i in 1:1000){
    alternative_data <- tibble(subjects = 1:(levels*n_per_level),
                          IV = as.factor(rep(1:levels, each = n_per_level)),
                          DV = c(rnorm(n_per_level, 125, 25),
                                 rnorm(n_per_level, 100, 25),
                                 rnorm(n_per_level, 100, 25),
                                 rnorm(n_per_level, 100, 25)
                                 )
                          )
    aov.out <- aov(DV ~ IV, data = alternative_data)
    simulated_F <- summary(aov.out)[[1]]$`F value`[1]
    save_altF_values[i] <- simulated_F
  }
  
  power <- length(save_altF_values[save_altF_values > qf(.95,3,(n_per_level*4)-4)])/1000
  sim_power[n] <- power
}

power_curve <- tibble(num_subjects,
                      sim_power)

knitr::kable(power_curve)
```

```{r}
ggplot(power_curve, aes(x=num_subjects,
                        y=sim_power))+
  geom_point()+
  geom_line()
```

```{r}
num_subjects <- c(10, 50, 75, 100, 125, 150, 175, 200)
sim_power <- length(length(num_subjects))

for(n in 1:length(num_subjects)){

  n_per_level <- num_subjects[n]
  
  # repeat the above many times to compute the F-distribution
  save_altF_values <- length(1000)
  for(i in 1:1000){
    alternative_data <- tibble(subjects = 1:(levels*n_per_level),
                          IV = as.factor(rep(1:levels, each = n_per_level)),
                          DV = c(rnorm(n_per_level, rnorm(1,0,.2), 1),
                                 rnorm(n_per_level, rnorm(1,0,.2), 1),
                                 rnorm(n_per_level, rnorm(1,0,.2), 1),
                                 rnorm(n_per_level, rnorm(1,0,.2), 1)
                                 )
                          )
    aov.out <- aov(DV ~ IV, data = alternative_data)
    simulated_F <- summary(aov.out)[[1]]$`F value`[1]
    save_altF_values[i] <- simulated_F
  }
  
  power <- length(save_altF_values[save_altF_values > qf(.95,3,(n_per_level*4)-4)])/1000
  sim_power[n] <- power
}

power_curve <- tibble(num_subjects,
                      sim_power)

knitr::kable(power_curve)
```

```{r}
ggplot(power_curve, aes(x=num_subjects,
                        y=sim_power))+
  geom_point()+
  geom_line()
```

```{r setup}
library(DBSStats2SemesterProject)
```
