---
title: "Lab7_FactorialAnova"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lab7_FactorialAnova}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Generalization Assignment

### 1. Explain the concept of main effects and interactions with an example using R. For example, this could include a definition of main effects and interactions and a figure depicting main effects and an interaction along with an explanation of the patterns for each. A major point of this problem is for you to to engage in the task of developing an explanation of these concepts that would 1) be helpful for you to understand the concepts, and 2) could be helpful for others to understand these concepts. (3 points)

```{r}

library(dplyr)
library(ggplot2)

water <- rep(c("No Water", "Water"), each = 20)
coffee <- rep(rep(c("No Coffee", "Coffee"), each = 10), 2)
subjects <- c(1:40)

caff_data <- c(sample(40:55, 10, replace = T), sample(30:40, 10, replace = T), sample(40:55, 10, replace = T), sample(65:85, 10, replace = T))

water_caff_data <- data.frame(subjects, water, coffee, caff_data)

water_caff_analysis <- summary(aov(caff_data ~ water*coffee, data = water_caff_data))
```

```{r}

### As the results of this 2x2 factorial ANOVA (hydration and caffeination levels as factors) on our fake data suggest, there are significant main effects of hydration levels (F[1, 36] = 190.05, MSE = 18, P < 0.001) and caffeination levels (F[1, 36] = 14.14, MSE = 18, P < 0.001), as well as a significant interaction effect between these two factors (F[1, 36] = 260.66, MSE = 18, P < 0.001.)

plot_caff <- water_caff_data %>%
            group_by(water, coffee) %>%
            summarise(mean_Energy = mean(caff_data),
                      SEM = sd(caff_data)/sqrt(length(caff_data)))

ggplot(plot_caff, aes(x=coffee, y=mean_Energy, group=water, fill=water))+
  geom_bar(stat="identity", position="dodge")+
  geom_errorbar(aes(ymin=mean_Energy-SEM, ymax=mean_Energy+SEM), 
                position=position_dodge(width=0.1),
                width=0.2)+
  theme_classic()+
  coord_cartesian(ylim=c(0,100))
```

```{r}

### This plot of our fake caffeine x hydration data visually confirms what can be interpreted from the ANOVA. Although There is little-to-no discernible difference between hydration conditions when the subjects have not had coffee, the main effect of hydration is likely driven by a substantial difference in energy levels between those that are hydrated and those that are not when they have had coffee. Similarly, the main effect of caffeination is likely driven by a substantial difference in energy levels between those that have or have not been hydrated. This explanation also underlies the highly significant interaction observed: The effects of one factor are dependent on the levels of the other factor. It could be described that coffee increased subjects' energy levels, but only when they were also hydrated. Conversely, it could be stated that hydration increased subjects' energy levels, but only when they were also caffeinated. 

### Struggle-report: I admit that I went to the solutions video for this, but in that video, the concept of main effects and interactions (which I was already familiar with) was the only thing explained for this question. So I pretty much did it on my own, but due to intent to check the solutions video, I'd rate myself maybe a 90-95% here. 

```

### 2. Complete the 2x2 factorial lab found here https://crumplab.github.io/statisticsLab/lab-10-factorial-anova.html, up to section 10.4.8. More specifically, your task is to follow that lab exercise to load in the data, transform the data into long-format, conduct a 2x2 between subjects ANOVA, and write a short results section reporting the main effects and interaction. (3 points)

```{r}

library(data.table)
stroop_data <- fread("https://raw.githubusercontent.com/CrumpLab/statisticsLab/master/data/stroop_stand.csv")

### I'm not able to get the "summarytools" packaged installed for some reason.

RTs <- c(as.numeric(unlist(stroop_data[,1])),
         as.numeric(unlist(stroop_data[,2])),
         as.numeric(unlist(stroop_data[,3])),
         as.numeric(unlist(stroop_data[,4]))
         )

Congruency <- rep(rep(c("Congruent","Incongruent"),each=50),2)
Posture <- rep(c("Stand","Sit"),each=100)
Subject <- rep(1:50,4)

stroop_df <- data.frame(Subject,Congruency,Posture,RTs)

library(tidyr)

stroop_long<- gather(stroop_data, key=Condition, value=RTs, 
                     congruent_stand, incongruent_stand,
                     congruent_sit, incongruent_sit)

new_columns <- tstrsplit(stroop_long$Condition, "_", names=c("Congruency","Posture"))

stroop_long <- cbind(stroop_long,new_columns)

stroop_long <- cbind(stroop_long,Subject=rep(1:50,4))

hist(stroop_long$RTs)

```

```{r}
library(dplyr)
library(ggplot2)

plot_means <- stroop_long %>%
            group_by(Congruency,Posture) %>%
            summarise(mean_RT = mean(RTs),
                      SEM = sd(RTs)/sqrt(length(RTs)))

ggplot(plot_means, aes(x=Posture, y=mean_RT, group=Congruency, fill=Congruency))+
  geom_bar(stat="identity", position="dodge")+
  geom_errorbar(aes(ymin=mean_RT-SEM, ymax=mean_RT+SEM), 
                position=position_dodge(width=0.9),
                width=.2)+
  theme_classic()+
  coord_cartesian(ylim=c(700,1000))
```

```{r}
library(xtable)

aov_out<-aov(RTs ~ Congruency*Posture, stroop_long)
summary_out<-summary(aov_out)

print(model.tables(aov_out,"means"), format="markdown")

library(xtable)
knitr::kable(xtable(summary_out))
```

```{r}
### The data of this postural-stroop experiment was subjected to a 2x2 between-subjects factorial ANOVA, with stimulus congruency and subject posture as factors. Mean response times for incongruent stimuli (922.3ms) were markedly slower than those for congruent stimuli (814.9ms), and this main effect was statistically significant (F[1, 196] = 43.73, MSE = 13189.185, P < 0.001). Mean response times differentiated according to postural condition were slightly different, with a mean time of 881.4ms for the "sitting" condition and 855.9ms for the "standing" condition. However, this difference did not achieve significance in our factorial ANOVA (F[1, 196] = 2.45, MSE = 13189.185, P = 0.1192). The interaction effect between the posture and congruency factors also failed to achieve significance (F[1, 196] = 0.49, MSE = 13189.185, P = 0.4815).

### Struggle-report: I was able to do this without watching the solutions video, but I was walked completely through by the original lab project itself. So I guess I technically got a 100% here, but it doesn't feel deserved. 
```

### 3. In chapter 10 of Crump et al. (2018), there is a discussion of patterns of main effects and interactions that can occur in a 2x2 design, which represents perhaps the simplest factorial design. There are 8 possible outcomes discussed https://crumplab.github.io/statistics/more-on-factorial-designs.html#looking-at-main-effects-and-interactions. Examples of these 8 outcomes are shown in two figures, one with bar graphs, and one with line graphs. Reproduce either of these figures using ggplot2. (3 points)

```{r}

g1 <- data.frame(IV1 = c("A", "A", "B", "B"),
                 IV2 = c("1", "2", "1", "2"),
                 means = c(5, 5, 5, 5))

g2 <- data.frame(IV1 = c("A", "A", "B", "B"),
                 IV2 = c("1", "2", "1", "2"),
                 means = c(5, 10, 10, 5))

g3 <- data.frame(IV1 = c("A", "A", "B", "B"),
                 IV2 = c("1", "2", "1", "2"),
                 means = c(5, 10, 5, 10))

g4 <- data.frame(IV1 = c("A", "A", "B", "B"),
                 IV2 = c("1", "2", "1", "2"),
                 means = c(2, 12, 5, 10))

g5 <- data.frame(IV1 = c("A", "A", "B", "B"),
                 IV2 = c("1", "2", "1", "2"),
                 means = c(10, 10, 5, 5))

g6 <- data.frame(IV1 = c("A", "A", "B", "B"),
                 IV2 = c("1", "2", "1", "2"),
                 means = c(10, 13, 5, 1))

g7 <- data.frame(IV1 = c("A", "A", "B", "B"),
                 IV2 = c("1", "2", "1", "2"),
                 means = c(5, 10, 10, 15))

g8 <- data.frame(IV1 = c("A", "A", "B", "B"),
                 IV2 = c("1", "2", "1", "2"),
                 means = c(10, 17, 5, 7))

all_plots <- rbind(g1, g2, g3, g4, g5, g6, g7, g8)

type <- c(rep("~1, ~2, ~1 x 2", 4),
          rep("1, ~2, ~1 x 2", 4),
          rep("~1, 2, 1 x 2", 4),
          rep("~1, 2, 1 x 2", 4),
          rep("1, ~2, ~1 x 2", 4),
          rep("1, ~2, 1 x 2", 4),
          rep("1, 2, ~1 x 2", 4),
          rep("1, 2, 1 x 2", 4))

type <- as.factor(type)

all_plots <- cbind(all_plots, type)

plots <- ggplot(all_plots, aes(x = IV1, y = means, group = IV2, fill = IV2))+
  geom_bar(stat = "identity", position = "dodge")+
  theme_classic()+
  facet_wrap(~type, nrow = 2)+
  theme(legend.position = "top")

```
```{r}
### I have no idea why it's only displaying 6 of the 8 graphs. 

### I needed to consult the solutions video to see where you were going with this, and ultimately needed to copy some of the prep in between the dataframe creation and the ggplot itself. However I refused to flat-out copy the numbers and formulae as you suggested in the solutions video. But because it was still so "copied" nevertheless, I'd only give myself maybe a 5-10% here.
```


### 4. In the conceptual section of this lab we used an R simulation to find the family-wise type I error rate for a simple factorial design with 2 independent variables. Use an R simulation to find the family-wise type I error rate for a factorial design with 3 independent variables. (3 points)

```{r}

save_sim <- tibble()

for(i in 1:10000){
  
  n <- 10
  factorial_data <- tibble(A = factor(rep(c("L1","L2"), each = n)),
                           B = factor(rep(c("A1","A2"), n)),
                           C = factor(rep(c("B1", "B2"), each = n)),
                           DV = rnorm(n*2,0,1))
  output <- summary(aov(DV~A*B*C, data=factorial_data))
  
  sim_tibble <- tibble(p_vals = output[[1]]$`Pr(>F)`[1:4],
                       effect = c("A","B", "C", "AxB"),
                       sim = rep(i,4))
  
  save_sim <-rbind(save_sim,sim_tibble)
}


type_I_errors <- save_sim %>%
  filter(p_vals < .05) %>%
  group_by(sim) %>%
  count()
  
dim(type_I_errors)[1]/10000

```

```{r}
### I'm not sure if I did this one right. But (copypasting and adapting from the in-class exercise component) I technically did not need any guidance from the solutions video. So 100%. 
```

### 5. Show that a 2x2 factorial ANOVA can be accomplished by using a one-factor ANOVA with three linear contrasts. (3 points)

```{r}

### I feel like skipping this one, I already technically did 2 extra anyway...and for this one I was probably going to copiously use the solutions video, in which you didn't actually explain this particular question.

```

## Follow-along with class assignment
```{r}

library(tibble)

n <- 10
factorial_data <- tibble(A = factor(rep(c("L1","L2"), each = n)),
                         B = factor(rep(c("L1","L2"), n)),
                         DV = rnorm(n*2,0,1))

```

```{r}

aov_out <- aov(DV ~ A*B, data = factorial_data)
summary(aov_out)
model.tables(aov_out, type = "means")

```

```{r}

n <- 12
factorial_data <- tibble(A = factor(rep(c("L1","L2"), each = n)),
                         B = factor(rep(rep(c("L1","L2"), each = n/2),2)),
                         C = factor(rep(c("L1","L2"), n)),
                         DV = rnorm(n*2,0,1))

summary(aov(DV ~ A*B*C, data = factorial_data))

```

```{r}

library(dplyr)
library(ggplot2)
library(patchwork)

n <- 10
factorial_data <- tibble(A = factor(rep(c("A1","A2"), each = n)),
                         B = factor(rep(c("B1","B2"), n)),
                         DV = rnorm(n*2,0,1))

factorial_data %>%
  group_by(A,B) %>%
  summarise(mean_DV = mean(DV))

factorial_data %>%
  ggplot(aes(y=DV, x=A, group = B,fill=B))+
   geom_bar(stat="summary", fun = "mean", position="dodge") +
   geom_point(position = position_dodge(width=0.5))

```

```{r}

A <- factorial_data %>%
  group_by(A) %>%
  summarise(mean_DV = mean(DV)) %>%
  ggplot(aes(y=mean_DV, x=A))+
   geom_bar(stat="identity", position="dodge") + 
   ggtitle("Main effect A")

B <- factorial_data %>%
  group_by(B) %>%
  summarise(mean_DV = mean(DV)) %>%
  ggplot(aes(y=mean_DV, x=B))+
   geom_bar(stat="identity", position="dodge")+ 
   ggtitle("Main effect B")

AB <- factorial_data %>%
  group_by(A,B) %>%
  summarise(mean_DV = mean(DV)) %>%
  ggplot(aes(y=mean_DV, x=A, fill=B))+
   geom_bar(stat="identity", position="dodge")+ 
   ggtitle("AxB Interaction")

(A+B)/AB

```

```{r}

aov_out <- aov(DV ~ A*B, data = factorial_data)
summary(aov_out)

model.tables(aov_out, type = "means")

```

```{r}

a1b1 <- c(11,9,7,11,12,7,12,11,10,10)
a1b2 <- c(12,12,7,9,9,10,12,10,7,12)
a2b1 <- c(13,18,19,13,8,15,13,9,8,14)
a2b2 <- c(13,21,20,15,17,14,13,14,16,7)
a3b1 <- c(17,20,22,13,21,16,23,19,20,19)
a3b2 <- c(32,31,27,30,29,30,33,25,25,28)

recall_data <- tibble(words_recalled = c(a1b1,a1b2,
                                         a2b1,a2b2,
                                         a3b1,a3b2),
                      A = rep(c("12 words",
                                "24 words",
                                "48 words"), each = 20),
                      B = rep(rep(c("Free recall",
                                "Cued Recall"), each = 10),3)
                      )

ggplot(recall_data, aes(x=A, y=words_recalled, group = B, linetype=B))+
  geom_point(stat="summary", fun="mean")+
  geom_line(stat="summary", fun="mean")

```

```{r}

aov_out <- aov(words_recalled ~ A*B, data = recall_data)

summary(aov_out)
model.tables(aov_out, type="means")

```

```{r}

A1 <- c(127,121,117,109,107,101,98,94,97,89)
A2 <- c(117,109,113,113,108,104,95,93,96,92)
A3 <- c(111,111,111,101,99,91,95,89,89,83)
A4 <- c(108,100,100,92,92,90,87,77,89,85)

random_data <- tibble(scores = c(A1,A2,A3,A4),
                      A = factor(rep(1:4,each = 10)),
                      B = factor(rep(rep(1:5,each=2),4))
                      )

aov.lm <- lm(formula = scores ~ A*B, data = random_data)
car::Anova(aov.lm, type = 2)

```

```{r}

aov_out <- aov(scores ~ A*B, data = random_data)
summary(aov_out)

```

```{r}

save_sim <- tibble()

for(i in 1:10000){
  
  n <- 10
  factorial_data <- tibble(A = factor(rep(c("L1","L2"), each = n)),
                           B = factor(rep(c("L1","L2"), n)),
                           DV = rnorm(n*2,0,1))
  output <- summary(aov(DV~A*B, data=factorial_data))
  
  sim_tibble <- tibble(p_vals = output[[1]]$`Pr(>F)`[1:3],
                       effect = c("A","B","AxB"),
                       sim = rep(i,3))
  
  save_sim <-rbind(save_sim,sim_tibble)
}

```

```{r}

type_I_errors <- save_sim %>%
  filter(p_vals < .05) %>%
  group_by(sim) %>%
  count()
  
dim(type_I_errors)[1]/10000

```

```{r}

save_sim %>%
  group_by(effect) %>%
  summarise(type_I_error = length(p_vals[p_vals < .05])/10000)

```

```{r}

a <- rbinom(10000,3,.05)
length(a[a>0])/10000

```

```{r setup}
library(DBSStats2SemesterProject)
```
