---
title: "Practicals_for_Final"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Practicals_for_Final}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Demonstrating types of objects

## Vectors, with averaging example

```{r}
library(dplyr)
vector_example <- c(4, 1, 9, 4, 2, 8, 7, 3)
mean(vector_example)
```

## Data frame, with the average for one level of a particular variable

```{r}

colors <- as.character(c(rep("blue", "red", "yellow", "purple", "green"), each = 2))

data_frame_example <- data.frame(colors = c("blue", "blue", "red", "red", "yellow", "yellow", "purple", "purple", "green", "green"),
                                congruency = c("congruent", "incongruent", "congruent", "incongruent", "congruent", "incongruent","congruent", "incongruent", "congruent", "incongruent"),
                                data = c(35, 49, 21, 92, 54, 17, 38, 43, 20, 31))
```

### Congruent mean
```{r}
average_congruent <- mean(data_frame_example$data[data_frame_example$congruency == "congruent"])
average_congruent
```

## List (utility exemplified with ice cream review example)
```{r}

names(bnj) <- c("flavor", "tastyrating", "creamyrating", "pricing")
bnj <- list("phishfood", 7, 4, "med")

hagendaaz <- list("vanilla", 6, 9, "high")
names(hagendaaz) <- c("flavor", "tastyrating", "creamyrating", "pricing")

breyers <- list("brownie", 9, 3, "low")
names(breyers) <- c("flavor", "tastyrating", "creamyrating", "pricing")

icecreamreviews <- list(bnj, hagendaaz, breyers)

### To conveniently vectorize tastyness ratings across two brands (the utility of lists is intended to be demonstrated here under a context of putting in all the review details of each ice cream brand separately)

hd_breyers_tasty_compare <- c(icecreamreviews[[2]]$tastyrating, icecreamreviews[[3]]$tastyrating)
hd_breyers_tasty_compare

```

## Matrix - Useful for turning a single vector into a table
```{r}

matrix_data <- c(rnorm(100, 0, 1))

matrix_example <- matrix(data = matrix_data, nrow = 10, ncol = 10, byrow = FALSE)

```

#### Matrices allow convenient triangulation of locations in your dataset- for example, here are two means:
```{r}
second_col_mean <- mean(matrix_example[,2])
sixth_row_mean <- mean(matrix_example[6,])

means_of_second_col_and_sixth_row <- c(second_col_mean, sixth_row_mean)
means_of_second_col_and_sixth_row
```

# Using a "for" loop 
## Demonstrating the obvious point that with a ratio (let's pretend we're talking about F), as the denominator ("within-group differences") gets larger, the overall ratio ("effect") gets smaller. Output gets smaller with each iteration.
```{r}

for(i in 0:20){
 pretend_F_ratio <- (rnorm(1, 0, 0.5)+1) / ((rnorm(1, 0, 0.5)^2) + i)
  print(pretend_F_ratio)
}

```

### Using a "while" loop for the same exact purpose; also demonstrating 

```{r}
breakpoint <- 20
while(i <= 20){
  pretend_F_ratio <- ((rnorm(1, 0, 0.5)^2)) / (rnorm(1, 0, 0.5)*i)
  if((pretend_F_ratio) < 0.05){
    break
  }
  print(pretend_F_ratio)
}

```

```{r setup}
library(DBSStats2SemesterProject)
```
